|

### 2022-08-19 paper/paper.Rmd

---
title: |
  New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions
type: Short Technical Note
author:
  - name: Alex Aumann
    affil: a
    email: aaum0002@student.monash.edu
  - name: German Valencia
    affil: a
    email: german.valencia@monash.edu
  - name: Ursula Laa
    affil: b
    email: ursula.laa@boku.ac.at
  - name: Dianne Cook
    affil: c
    email: dicook@monash.edu
affiliation:
  - num: a
    address: |
      School of Physics and Astronomy, Monash University
  - num: b
    address: |
      Institute of Statistics, University of Natural Resources and Life Sciences, Vienna
  - num: b
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
abstract: |
  Something here
keywords: |
  data visualisation; grand tour; statistical computing; statistical graphics; multivariate data; dynamic graphics
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
  \usepackage{graphicx}
  \usepackage{nicematrix}
  \NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}
output: rticles::tf_article
keep_tex: true
---

@@

@@

>>>>

---
title: |
  New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions
type: Short Technical Note
author:
  - name: Ursula Laa
    affil: a
    email: ursula.laa@boku.ac.at
  - name: Alex Aumann
    affil: b
    email: aaum0002@student.monash.edu
  - name: Dianne Cook
    affil: c
    email: dicook@monash.edu
  - name: German Valencia
    affil: b
    email: german.valencia@monash.edu
affiliation:
  - num: a
    address: |
      Institute of Statistics, University of Natural Resources and Life Sciences, Vienna
  - num: b
    address: |
      School of Physics and Astronomy, Monash University
  - num: c
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
abstract: | 
  We implement a new version of a manual tour that takes advantage of existing features in Mathematica to construct functions that interact with the position of the cursor. We demonstrate the usefulness of this new tool with an example exploring region boundaries in classification problems.
keywords: |
  data visualisation; grand tour; statistical computing; statistical graphics; multivariate data; dynamic graphics
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
  \usepackage{graphicx}
  \usepackage{nicematrix}
  \NiceMatrixOptions{code-for-first-row = \color{red} ,code-for-last-row = \color{red} ,code-for-first-col = \color{blue} ,code-for-last-col = \color{blue}}
output: rticles::tf_article
keep_tex: true
---

@@

{\textcolor{blue}Another very useful feature is that the user can keep the projection fixed while slicing across the orthogonal space and changing the centre point used for slicing. This can serve to illustrate how the orthogonal space affects correlations in a slice and how they may change or disappear in a projection.}

@@

Mathematica provides much utility and versatility, such as inbuilt data visualisation, data manipulation and analysis, dynamic functionality, and symbolic and numeric computation. Most of this inbuilt functionality is quite user-friendly and can be found in the Mathematica documentation; typically, numerous examples are provided within the documentation, and some possible issues are outlined there. Also, it is relatively simple  to create dynamic objects via the inbuilt commands Manipulate or DynamicModule. Control objects, such as sliders, locator panes, and input fields, can then be used on dynamic variables, and when there is a change in the dynamic variable the dynamic objects which contain that dynamic variable will be updated. This is the essential ingredient for this implementation of the manual tour. However, it can be hard to alter the behaviour of the control objects such as setting the 1D slider on a different scale  (Slider increments add 10 points into the slice rather than increasing the slice height by 0.1 for example).
{\textcolor{red}Alex, is this still true with the controls you are using now?}

The examples presented in the penguins’ notebooks all utilise our primary new function, SliceDynamic. This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index. After specifying the initial slice height and the slice range, the user is then presented with an interactive display in which the control objects appear on the left and the slice visulisation appears on the right. The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice height; and there is an input field, which changes the slice centre. The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided. The current projection can be displayed to contrast it with the slice by ticking a box. The explicit projection matrix can be displayed via another checkbox. The Slice Tour* notebooks include the additional functions: ProjectionPlot, Projected2DSliderPlot, and VisualiseSliceDyanmic. ProjectionPlot is very similar to SliceDynamic, except it only displays projections and VisualiseSliceDynamic allows you to discern which points exist inside the slice and outside it. However, ProjectedLocatorPlot displays the interactive slightly differently. Several locator panes are displayed above the plot and each one corresponds to a row in the projection matrix but behaves the same as ProjectionPlot. {\textcolor{blue}(this may be easier to implement in another language (?) )}
